---
params:
  circuit: 'Heat Pump'
  title: 'GREENGrid Household Electricity Demand Data Circuit Extract Test: '
title: '`r paste0(params$title, params$circuit)`'
subtitle: '`r paste0(params$subtitle, params$hhID)`'
author: "Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    toc: yes
    toc_depth: 2
    toc_float: TRUE
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
bibliography: '`r path.expand("~/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
```


```{r codeSetup, include=FALSE}

# Set start time ----
startTime <- proc.time()

# Local parameters ----
library(GREENGrid)
GREENGrid::setup() # set package parameters mostly from ggParams.R

# fix to suit
gsFile <- path.expand("~/Dropbox/Work/Otago_CfS_Ben/data/nzGREENGrid/dataExtracts/Heat Pump_2015-04-01_2016-03-31_observations.csv.gz")
hhFile <-  path.expand("~/Dropbox/Work/Otago_CfS_Ben/data/nzGREENGrid/ggHouseholdAttributesSafe.csv")
  
# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "lubridate", # date & time stuff
             "ggplot2", # for fancy graphs
             "readr", # for reading & parsing .csv files
             "rmarkdown", # for render
             "bookdown", # for html2
             "GREENGridData", # data loading functions
             "kableExtra" # for pretty kable
)
# load them
GREENGrid::loadLibraries(rmdLibs)

# Local functions ----


```

\newpage

# About

## Report circulation:

 * Public - analysis for use in ICEERB 2018 presentation.
 
## License

```{r ccby license, child=ggParams$licenseCCBY}
```
 
## Citation

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r lubridate::year(today())`) `r paste0(params$title,params$circuit)`. `r ggParams$pubLoc`.

This work is (c) `r lubridate::year(today())` the University of Southampton.

## History

 * [Report history](https://github.com/CfSOtago/GREENGrid/commits/master/analysis/heatPumpAnalysis/heatPumpAnalysis.Rmd)
 
## Requirements:

This report uses the safe version of the grid spy 1 minute data which has been processed using the code in https://github.com/CfSOtago/GREENGridData/tree/master/dataProcessing/gridSpy. It also assumes you have already run the example circuit extraction script using circuit = `r params$circuit`.

## Support

```{r generic support, child=ggParams$supportGeneric}
```
 
\newpage

# Introduction

```{r generic sample, child=ggParams$sampleGeneric}
```
 
This report provides summary analysis of one circuit type (`r params$circuit`) across all households as an example.

# Load data

The data used to generate this report is:

 * `r gsFile`
 * `r hhFile`

First we load the household data. `readr` will give some feedback on the columns.

```{r loadHhData}
if(file.exists(hhFile)){
  hhDT <- data.table::as.data.table(readr::read_csv(hhFile, progress = FALSE)) # load hh data
  data.table::setkey(hhDT, linkID)
} else {
  print(paste0("Failed to find ", hhFile," - is the data source available?"))
  } 

```

Next we load the Grid Spy extract for `r params$circuit`. This uses a `GREENGridData` package function intended to load the cleaned individual household data which warns that two of the column names are not found. These columns were dropped during the extraction process so we can safely ignore these warnings

```{r loadGsData}
gsDT <- GREENGridData::getCleanGridSpyFile(gsFile) # load Grid Spy data
data.table::setkey(gsDT, linkID)
t <- head(gsDT)

kableExtra::kable(t, caption = paste0("First few rows of grid spy data")) %>%
  kable_styling()
```

Table \@ref(tab:loadGsData) shows the first few rows of the Grid Spy 1 minute power data. 

```{r summaryGsData}
t <- summary(gsDT)

kableExtra::kable(t, caption = paste0("Summary of grid spy data")) %>%
  kable_styling()

# create some useful derived date & time variables
gsDT <- gsDT[, obsTime := hms::as.hms(r_dateTime)] # HH:MM for demand profile plots

```

Table \@ref(tab:summaryGsData) shows a summary of the Grid Spy 1 minute power data. 

Note that we have some Nega watts - which households have them?

```{r negaWatts}
# we've got negawatts - filter them out
gsDT <- gsDT[, negW := "PosW"]
gsDT <- gsDT[ powerW < 0, negW := "NegaW"]

t <- table(gsDT$linkID,gsDT$negW)
t
round(100*prop.table(t,1),1)

gsDT <- gsDT[negW == "PosW"] # remove negawatts

t <- summary(gsDT)

kableExtra::kable(t, caption = paste0("Summary of cleaned grid spy data (check for NAs)")) %>%
  kable_styling()
```

Table \@ref(tab:negaWatts) shows a summary of the Grid Spy 1 minute power data after the removal of any negaWatts. 

Note that:

 * r_datetime is the correct dateTime of each observation in UTC and will have loaded as your local timezone. If you are conducting this analysis outside NZ then you will get strange results until you use [lubridate](https://lubridate.tidyverse.org/) to tell R to use tz = "Pacific/Auckland" with this variable;
 * there can be 0 Wh observations.

# Plot seasonal mean power profiles

First we create a _Southern_ Hemisphere season variable. Luckily we have a function to do this in the `GREENGridData` package. We print a check table to ensure we are all happy with the coding of `season`.

```{r addSeason, echo = TRUE}
gsDT <- GREENGridData::addNZSeason(gsDT)
table(lubridate::month(gsDT$r_dateTime, label = TRUE), gsDT$season, useNA = "always")
```

For simplicity we will focus only on Summer and Winter.

```{r subsetSeasons}
gsDT <- gsDT[season == "Winter" | season == "Summer"]
```

## Overall profiles

This section plots overall mean power per half hour by season.

```{r makePlot, fig.cap="Demand profile plot", fig.height=6, echo = TRUE}
gsDT <- gsDT[, r_dateTimeQHour := lubridate::floor_date(r_dateTime, unit = "15 mins")]

# create mean power across 15 minute periods to use as base dataset (comparable to SAVE)
qHourDT <- gsDT[, .(meanW = mean(powerW)), keyby = .(r_dateTimeQHour,linkID, season)
             ]
qHourDT <- qHourDT[, obsQHour := hms::as.hms(r_dateTimeQHour)]

plotDT <- qHourDT[, .(meanW = mean(meanW)), keyby = .(season, obsQHour)
             ]

# set attributes for plot
vLineAlpha <- 0.4
vLineCol <- "#0072B2" # http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette
timeBreaks <- c(hms::as.hms("04:00:00"), 
                hms::as.hms("08:00:00"),
                hms::as.hms("12:00:00"),
                hms::as.hms("16:00:00"),
                hms::as.hms("20:00:00"),
                hms::as.hms("24:00:00")
)

# create default caption
myCaption <- paste0("GREENGrid Grid Spy household electricity demand data (https://dx.doi.org/10.5255/UKDA-SN-853334)",
                        "\n", min(lubridate::date(gsDT$r_dateTime)), 
                        " to ", max(lubridate::date(gsDT$r_dateTime)),
                        "\nTime = Pacific/Auckland",
                        "\n (c) ", lubridate::year(now())," University of Otago")

myPlot <- ggplot2::ggplot(plotDT[!is.na(season)], # make sure no un-set seasons/non-parsed dates
                          aes(x = obsQHour, y = meanW/1000)) +
  geom_line() + 
  facet_grid(season ~ .) +
  scale_colour_manual(values=ggParams$cbPalette) + # use colour-blind friendly palette
  theme(strip.text.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + 
  guides(colour = guide_legend(title = "Season: ")) +
  theme(legend.position = "bottom")  + 
  labs(title = paste0(params$circuit, ": seasonal mean power demand profiles"),
       y = "Mean kW per 15 minutes", 
       x = "Time of day",
       caption = myCaption
       )

myPlot + 
  scale_x_time(breaks = timeBreaks) +
  geom_vline(xintercept = timeBreaks, alpha = vLineAlpha, colour = vLineCol)

#ggplot2::ggsave(paste0(ggParams$repoLoc,"/examples/outputs/", params$circuit, "_meankWperminBySeason.png"))
```

Figure \@ref(fig:makePlot) shows the overall mean kW per minute in each season for this circuit (`r params$circuit`).

## Profiles by linked household attributes

### Number of people

Table \@ref(tab:hhPeopleTable) shows the number of households who have different numbers of people (children and adults). This table includes households where we do not know the number of people (NA) but we do have electricity demand data.

```{r hhPeopleTable}
t <- hhDT[, .(Freq = .N), keyby = Q57]

kableExtra::kable(t, caption = paste0("Number of households with number of persons")) %>%
  kable_styling()
```

Clearly this is too fine grained (too many categories). We therefor collapse to form the coding shown in \@ref(tab:hhPeopleTableReduced).

```{r hhPeopleTableReduced}
hhDT <- hhDT[Q57 == 1, nPeople := "1"]
hhDT <- hhDT[Q57 == 2, nPeople := "2"]
hhDT <- hhDT[Q57 == 3, nPeople := "3"]
hhDT <- hhDT[Q57 > 3, nPeople := "4+"]

t <- hhDT[, .(Freq = .N), keyby = nPeople]

kableExtra::kable(t, caption = paste0("Number of households with number of persons (recoded)")) %>%
  kable_styling()
```

Now we link (join) the Grid Spy and household data.tables and aggregate (summarise) by season and number of people. You can do this using `data.table`'s on the fly [join](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-faq.html#MergeDiff) but we have found pre-joining of the columns you want to be much faster. We're not sure why as it shouldn't be. You can probably also do this in `dplyr` etc but we haven't tried. 

Figure \@ref(fig:nPeoplePlot) shows the mean kW per minute per season by presence of young children for this circuit (`r params$circuit`). Can you see anything interesting or unusual and might this be due to the numbers of households in each group?

```{r nPeoplePlot, fig.height=6, fig.cap="Demand profile plot - n people"}
keepCols <- c("linkID", "nPeople")
setkey(qHourDT, linkID)
mergedDT <- qHourDT[hhDT[, ..keepCols]]
plotDT <- mergedDT[!is.na(nPeople), .(meanW = mean(meanW),
                                      sdW = sd(meanW),
                                      nObs = .N), keyby = .(season, obsQHour, nPeople)]
plotDT <- plotDT[, ci_upper := meanW + qnorm(0.975)*(sdW/sqrt(nObs))]
plotDT <- plotDT[, ci_lower := meanW - qnorm(0.975)*(sdW/sqrt(nObs))]

basePlot <- ggplot2::ggplot(plotDT[!is.na(season)], # make sure no un-set seasons/non-parsed dates
                          aes(x = obsQHour, y = meanW/1000, 
                              colour = nPeople)) +
  geom_line() + 
  scale_colour_manual(values=ggParams$cbPalette) + # use colour-blind friendly palette
  facet_grid(season  ~ .) + 
  theme(strip.text.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + 
  guides(colour = guide_legend(title = "Number of people: ")) +
  theme(legend.position = "bottom")  + 
  labs(title = paste0(params$circuit, ": seasonal mean power demand profiles by n people"),
       y = "Mean kW per minute",
       x = "Time of day",
       caption = myCaption
       )

basePlot <- basePlot + 
  scale_x_time(breaks = timeBreaks) +
  geom_vline(xintercept = timeBreaks, alpha = vLineAlpha, colour = vLineCol)

basePlot

# add 95% CI

ciPlot <- basePlot + geom_errorbar(aes(ymin = ci_lower/1000, ymax = ci_upper/1000))
ciPlot
```

### Number of children

This section plots overall mean power per minute by season and number of children aged 0-12 as an illustration of how to link the Grid Spy and household data. We will go through the steps with commentary and showing the code...

Table \@ref(tab:hhKidsTable) shows the number of households who have different numbers of children aged 0-12 so we know how many households make up each line on the plot. This table includes households where we do not know the number of children (NA) but we do have electricity demand data.

```{r hhKidsTable}
t <- hhDT[, .(Freq = .N), keyby = nChildren0_12]

kableExtra::kable(t, caption = paste0("Number of households with children aged 0-12")) %>%
  kable_styling()
```


Now use the aggregated `data.table` to make the plot. Note that as specified this will add a line for nChildren0_12 == NA household(s) - see Table \@ref(tab:hhTable).

```{r makeKidsPlot, fig.height=6, fig.cap="Demand profile plot - n kids", echo = TRUE}
keepCols <- c("linkID", "nChildren0_12")
mergedDT <- qHourDT[hhDT[, ..keepCols]]
plotDT <- mergedDT[!is.na(nChildren0_12), .(meanW = mean(meanW),
                                      sdW = sd(meanW),
                                      nObs = .N), keyby = .(season, obsQHour, nChildren0_12)]

plotDT <- plotDT[, ci_upper := meanW + qnorm(0.975)*(sdW/sqrt(nObs))]
plotDT <- plotDT[, ci_lower := meanW - qnorm(0.975)*(sdW/sqrt(nObs))]

basePlot <- ggplot2::ggplot(plotDT[!is.na(season)], # make sure no un-set seasons/non-parsed dates
                          aes(x = obsQHour, y = meanW/1000, 
                              colour = as.factor(nChildren0_12))) +
  geom_line() + 
  scale_colour_manual(values=ggParams$cbPalette) + # use colour-blind friendly palette
  facet_grid(season  ~ .) + 
  theme(strip.text.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + 
  guides(colour = guide_legend(title = "Number of children aged 0 - 12: ")) +
  theme(legend.position = "bottom")  + 
  labs(title = paste0(params$circuit, ": seasonal mean power demand profiles by n children aged 0-12"),
       y = "Mean kW per 15 minutes",
       x = "Time of day",
       caption = myCaption
       )

basePlot <- basePlot + 
  scale_x_time(breaks = timeBreaks) +
  geom_vline(xintercept = timeBreaks, alpha = vLineAlpha, colour = vLineCol)

basePlot

# add 95% CI

ciPlot <- basePlot + geom_errorbar(aes(ymin = ci_lower/1000, ymax = ci_upper/1000))
ciPlot
```

Figure \@ref(fig:makeKidsPlot) shows the mean kW per minute per season by presence of young children for this circuit (`r params$circuit`). Can you see anything interesting or unusual and might this be due to the numbers of households in each group?

# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

## R packages used

 * base R [@baseR]
 * bookdown [@bookdown]
 * GREENGridData [@GREENGridData] which depends on:
    - data.table [@data.table]
    - dplyr [@dplyr]
    - hms [@hms]
    - lubridate [@lubridate]
    - progress [@progress]
    - readr [@readr]
    - readxl [@readxl]
    - reshape2 [@reshape2]
 * kableExtra [@kableExtra]
 * knitr [@knitr]
 * rmarkdown [@rmarkdown]

## Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
