---
title: 'Processing, cleaning and saving NZ GREEN Grid project 1 minute electricity
  consumption data'
author: 'Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)'
date: 'Last run at: `r Sys.time()`'
output:
  html_document:
    code_folding: hide
    fig_caption: true
    keep_md: true
    number_sections: true
    self_contained: no
    toc: true
    toc_float: true
    toc_depth: 2
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r codeSetup, include=FALSE}
# Housekeeping ----
rm(list=ls(all=TRUE)) # remove all objects from workspace

# Set start time ----
startTime <- proc.time()

# Load greenGrid package ----
library(greenGridr) # local utilities

# Libraries ----
reqLibsLocal <- c("data.table", # for data munching
                  "lubridate", # for date/time munching
                  "ggplot2", # for fancy graphs
                  "readr", # for read/write_csv
                  "dplyr", # for select columns
                  #"dtplyr", # if needed
                  "knitr" # for kable
                  )

myRequiredPackages(reqLibsLocal)

# Local parameters ----
fullFb <- 0 # switch on full feedback
baTest <- 1 # test or full run?

if(baTest == 1){
  # Local test
  dPath <- "~/Data/NZGreenGrid/gridspy/" # BA laptop test set
  fpath <- paste0(dPath,"1min_orig/") # location of data
  outPath <- paste0(dPath, "consolidated/") # place to save them - add "1min/" for folder etc
} else {
  # full monty
  dPath <- "/Volumes/hum-csafe/Research Projects/GREEN Grid/" # HPS
  fpath <- paste0(dPath,"_RAW DATA/GridSpyData/") # location of data
  outPath <- paste0(dPath, "Clean_data/gridSpy/") # place to save them - add "1min/" for folder etc
}

pattern1Min <- "*at1.csv$" # e.g. *at1.csv$ filters only 1 min data
indexFile <- "fListCompleteDT.csv" # place to store the complete file list with meta-data
dataThreshold <- 3000 # assume any files smaller than this (bytes) = no data or some mangled xml/html. Really, we should check the contents of each file.

```

\newpage

# Citation

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r 1900 + as.POSIXlt(Sys.Date())$year`) Processing, cleaning and saving NZ GREEN Grid project 1 minute electricity consumption data, University of Otago: Dunedin, NZ.

\newpage

# Introduction

Report circulation:

 * Restricted to: [NZ GREEn Grid](https://www.otago.ac.nz/centre-sustainability/research/energy/otago050285.html) project partners and contractors.

## Purpose

This report is intended to: 

 * load and clean the project electricity consumption data (Grid Spy)
 * save the cleaned data out as a single file per household
 * produce summary data quality statistics

## Requirements:

 * grid spy 1 minute data downloads

## History

Generally tracked via our git.soton [repo](https://git.soton.ac.uk/ba1e12/nzGREENGrid).
 
## Support

This work was supported by:

 * The [University of Otago](https://www.otago.ac.nz/)
 * The New Zealand [Ministry of Business, Innovation and Employment (MBIE)](http://www.mbie.govt.nz/)
 * [SPATIALEC](http://www.energy.soton.ac.uk/tag/spatialec/) - a [Marie Skłodowska-Curie Global Fellowship](http://ec.europa.eu/research/mariecurieactions/about-msca/actions/if/index_en.htm) based at the University of Otago’s [Centre for Sustainability](http://www.otago.ac.nz/centre-sustainability/staff/otago673896.html) (2017-2019) & the University of Southampton's Sustainable Energy Research Group (2019-202).
 
This work uis (c) `r as.POSIXlt(Sys.time())$year + 1900` the University of Southampton.

# Obtain listing of files

In this section we generate a listing of all 1 minute data files that we have received. If we are running over the complete dataset then we will be using data from:

 * /hum-csafe/Research Projects/GREEN Grid/_RAW DATA/GridSpyData/
 
In this run we are using data from:

 * `r fpath`

If these do not match then this may be a test run.

```{r getCompleteFileList}

print(paste0("Looking for 1 minute data using pattern = ", pattern1Min, " in ", fpath))
fListCompleteDT <- as.data.table(list.files(path = fpath, pattern = pattern1Min, # use the default pattern to filter e.g. 1m from 30s files
                                            recursive = TRUE))
if(nrow(fListCompleteDT) == 0){
  stop(paste0("No matching data files found, please check your path (", fpath, ") or your search pattern (", pattern1Min, ")"))
} else {
  print(paste0("Processing file list and getting file meta-data (please be patient)"))
  fListCompleteDT <- fListCompleteDT[, c("hhID","fileName") := tstrsplit(V1, "/")]
  fListCompleteDT <- fListCompleteDT[, fullPath := paste0(fpath, hhID,"/",fileName)]
  
  # now loop over the files and collect metadata
  for(f in fListCompleteDT[,fullPath]){
    if(baTest | fullFb){print(paste0("Checking ", f))}
    rf <- path.expand(f) # just in case of ~ etc
    fsize <- file.size(rf)
    fmtime <- ymd_hms(file.mtime(rf), tz = "Pacific/Auckland") # requires lubridate
    fListCompleteDT <- fListCompleteDT[fullPath == f, fSize := fsize]
    fListCompleteDT <- fListCompleteDT[fullPath == f, fMTime := fmtime]
    fListCompleteDT <- fListCompleteDT[fullPath == f, fMDate := as.Date(fmtime)]
    fListCompleteDT <- fListCompleteDT[fullPath == f, dateColName := "unknown - not loaded"]
    # only try to read files where we think there might be data
    if(fsize > dataThreshold){
      if(fullFb){print(paste0("fSize (", fsize, ") > threshold (", dataThreshold, ") -> loading ", f))}
      row1DT <- fread(f, nrows = 1)
      # what is the date column called?
      fListCompleteDT <- fListCompleteDT[fullPath == f, dateColName := "unknown - can't tell"]
      if(nrow(select(row1DT, contains("NZ"))) > 0){ # requires dplyr
        setnames(row1DT, 'date NZ', "dateTime_char")
        fListCompleteDT <- fListCompleteDT[fullPath == f, dateColName := "date NZ"]
      } 
      if(nrow(select(row1DT, contains("UTC"))) > 0){ # requires dplyr
        setnames(row1DT, 'date UTC', "dateTime_char")
        fListCompleteDT <- fListCompleteDT[fullPath == f, dateColName := "date UTC"]
      }
      # split dateTime
      row1DT <- row1DT[, c("date_char", "time_char") := tstrsplit(dateTime_char, " ")]
      # add example of date to metadata - presumably they are the same in each file?!
      fListCompleteDT <- fListCompleteDT[fullPath == f, dateExample := row1DT[1, date_char]]
      
      if(fullFb){print(paste0("Checking date formats in ", f))}
      row1DT <- checkDates(row1DT)
      fListCompleteDT <- fListCompleteDT[fullPath == f, dateFormat := row1DT[1, dateFormat]]
      if(fullFb){print(paste0("Done ", f))}
    }
  }
  if(baTest | fullFb){print("All files checked")}
  # any date formats are still ambiguous need a deeper inspection using the full file - could be slow
  fAmbig <- fListCompleteDT[dateFormat == "ambiguous", fullPath]
  
  for(fa in fAmbig){
    if(baTest | fullFb){print(paste0("Checking ambiguous date formats in ", fa))}
    dt <- fread(fa)
    if(nrow(select(dt, contains("NZ"))) > 0){ # requires dplyr
      setnames(dt, 'date NZ', "dateTime_char")
      
    } 
    if(nrow(select(dt, contains("UTC"))) > 0){ # requires dplyr
      setnames(dt, 'date UTC', "dateTime_char")
      
    }
    dt <- dt[, c("date_char", "time_char") := tstrsplit(dateTime_char, " ")]
    dt <- checkDates(dt)
    # set what we now know (or guess!)
    fListCompleteDT <- fListCompleteDT[fullPath == f, dateFormat := dt[1,dateFormat]]
  }
      
  ofile <- paste0(outPath, indexFile)
  print(paste0("Saving 1 minute data files metadata to ", ofile))
  write.csv(fListCompleteDT, ofile)
  print("Done")
}

print(paste0("Overall we have ", nrow(fListCompleteDT), " files from ", uniqueN(fListCompleteDT$hhID), " households."))

t <- fListCompleteDT[, .(nFiles = .N), keyby = .(dateColName, dateFormat)]

kable(caption = "Number of files with given data column names", t)
```

Overall we have `r nrow(fListCompleteDT)` files from `r uniqueN(fListCompleteDT$hhID)` households. The following chart shows the distribution of these files over time using their sizes. Note that white indicates the presence of small files which may not contain observations.

```{r allFileSizesPlot}
myCaption <- paste0("Data source: ", fpath,
                    "\nUsing data received up to ", Sys.Date())

plotDT <- fListCompleteDT[, .(nFiles = .N,
                              meanfSize = mean(fSize)), 
                          keyby = .(hhID, date = as.Date(fMDate))]

#>> All files plots ----
ggplot(plotDT, aes( x = date, y = hhID, fill = log(meanfSize))) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "black") + 
  scale_x_date(date_labels = "%Y %b", date_breaks = "1 month") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + 
  labs(title = "Mean file size of all grid spy data files received per day",
       caption = paste0(myCaption, 
                        "\nLog file size used as some files are full year data")
    
  )
ggsave(paste0(outPath, "gridSpyAllFileListSizeTilePlot.png"))
```

# Load data files

In this section we load the data files that have a file size > `r dataThreshold` bytes. Things to note:

 * We assume that any files smaller than this value have no observations. This is based on:
     * Manual inspection of several small files
     * The identical (small) file sizes involved
     * _But_ we should probably test the first few lines to double check...
 * We have to deal with at least 2 different date formats and quite a lot of duplication some of which causes the differnt date formats. The code to fix these is included below but see also our [repo issues list](https://git.soton.ac.uk/ba1e12/nzGREENGrid/issues?scope=all&utf8=%E2%9C%93&state=all).
 

```{r loadValidFiles}
# > Load, process & save the ones which probably have data ----

hhIDs <- unique(fListCompleteDT$hhID) # list of household ids
hhStatDT <- data.table() # stats collector
for(hh in hhIDs){
  print(paste0("Loading: ", hh))
  tempHhDT <- data.table() # create data.table to hold file contents
  fListCompleteDT <- fListCompleteDT[, fileLoaded := "No"]
  filesToLoad <- fListCompleteDT[hhID == hh & fSize > dataThreshold, fullPath] # select the files that meet our size threshold
  for(f in filesToLoad){
    if(baTest){print(paste0("File size (", f, ") = ", 
                            fListCompleteDT[fullPath == f, fSize], 
                            " so probably OK"))} # files under 3kb are probably empty
    # attempt to load the file
    tempDT <- fread(f)
    if(baTest){print("File loaded")}
    # set some file stats
    fListCompleteDT <- fListCompleteDT[fullPath == f, fileLoaded := "Yes"]
    fListCompleteDT <- fListCompleteDT[fullPath == f, nObs := nrow(tempDT)] # could include duplicates
    
    if(nrow(select(tempDT, contains("NZ"))) > 0){ # requires dplyr
      # => there is at least 1 column whose name contains NZ so we have NZ time
      if(baTest){print("NZ time")}
      setnames(tempDT, 'date NZ', "date_NZ")
      # Check the date format as it could be y-m-d or d/m/y or m/d/y :-(
      tempDT <- tempDT[, c("date_char", "time_char") := tstrsplit(date_NZ, " ")]
      try(tempDT <- tempDT[, c("date_char1", "date_char2", "date_char3") := tstrsplit(date_char, "/")])
      try(tempDT <- tempDT[, c("date_char1", "date_char2", "date_char3") := tstrsplit(date_char, "-")])
      tempDT$dateFormat <- "ambiguous"
      if(max(as.integer(tempDT$date_char3)) > 32){
        # char 3 = year
        if(max(as.integer(tempDT$date_char2)) > 12){
          # char 2 = day
          tempDT$dateFormat <- "mdy"
        }
        if(max(as.integer(tempDT$date_char1)) > 12){
          # char 1 = month
          tempDT$dateFormat <- "dmy"
        }
      }
      if(max(as.integer(tempDT$date_char1)) > 32){
        # char 1 = year
        if(max(as.integer(tempDT$date_char2)) > 12){
          # char 2 = day
          tempDT$dateFormat <- "ydm"
        }
        if(max(as.integer(tempDT$date_char1)) > 12){
          # char 1 = month
          tempDT$dateFormat <- "ymd"
        }
      }
      # Now use that to correctly parse dates
      tempDT <- tempDT[dateFormat == "mdy", r_dateTime := mdy_hm(date_NZ, tz = "Pacific/Auckland")] # requires lubridate
      tempDT <- tempDT[dateFormat == "dmy", r_dateTime := dmy_hm(date_NZ, tz = "Pacific/Auckland")] # requires lubridate
      tempDT <- tempDT[dateFormat == "ydm", r_dateTime := ymd_hm(date_NZ, tz = "Pacific/Auckland")] # requires lubridate
      tempDT <- tempDT[dateFormat == "ymd", r_dateTime := ymd_hm(date_NZ, tz = "Pacific/Auckland")] # requires lubridate
      if(baTest){
        print(head(tempDT))
        print(summary(tempDT))
        print(table(tempDT$dateFormat))
        }
    } else {
      # we have UTC
      if(baTest){print("UTC")}
      setnames(tempDT, 'date UTC', "date_UTC")
      tempDT <- tempDT[, c("date_char", "time_char") := tstrsplit(date_NZ, " ")]
      tempDT <- tempDT[, c("date_char1", "date_char2", "date_char3") := tstrsplit(date_char, "-")]
      tempDT$dateFormat <- "ambiguous"
      if(max(as.integer(tempDT$date_char3)) > 32){
        # char 3 = year
        if(max(as.integer(tempDT$date_char2)) > 12){
          # char 2 = day
          dateFormat <- "mdy"
        }
        if(max(as.integer(tempDT$date_char1)) > 12){
          # char 1 = month
          dateFormat <- "dmy"
        }
      }
      if(max(as.integer(tempDT$date_char1)) > 32){
        # char 1 = year
        if(max(as.integer(tempDT$date_char2)) > 12){
          # char 2 = day
          dateFormat <- "ydm"
        }
        if(max(as.integer(tempDT$date_char1)) > 12){
          # char 1 = month
          dateFormat <- "ymd"
        }
      }
      # Now use that to correctly parse dates
      tempDT$dateFormat <- dateFormat
      fListCompleteDT <- fListCompleteDT[fullPath == f, dateFormat := dateFormat]
      tempDT <- tempDT[dateFormat == "mdy", r_dateTime := mdy_hm(date_NZ, tz = "UTC")] # requires lubridate
      tempDT <- tempDT[dateFormat == "dmy", r_dateTime := dmy_hm(date_NZ, tz = "UTC")] # requires lubridate
      tempDT <- tempDT[dateFormat == "ydm", r_dateTime := ymd_hm(date_NZ, tz = "UTC")] # requires lubridate
      tempDT <- tempDT[dateFormat == "ymd", r_dateTime := ymd_hm(date_NZ, tz = "UTC")] # requires lubridate
      if(baTest){
        print(head(tempDT))
        print(summary(tempDT))
        print(table(tempDT$dateFormat))
        }
    }
    fListCompleteDT <- fListCompleteDT[fullPath == f, obsStartDate := min(as.Date(tempDT$r_dateTime))]
    fListCompleteDT <- fListCompleteDT[fullPath == f, obsEndDate := max(as.Date(tempDT$r_dateTime))]
    tempHhDT <- rbind(tempHhDT, tempDT, fill = TRUE) # just in case there are different numbers of columns (quite likely!)
  }
  
  # > Remove duplicates caused by over-lapping files and dates etc ----
  # Need to remove all test vars for this
  try(tempHhDT$date_UTC <- NULL)
  try(tempHhDT$date_NZ <- NULL)
  try(tempHhDT$testDate <- NULL)
  
  #nObs <- nrow(tempHhDT)
  #print(paste0("N rows before removal of duplicates: ", nObs))
  tempHhDT <- unique(tempHhDT)
  #nObs <- nrow(tempHhDT)
  #print(paste0("N rows after removal of duplicates: ", nObs))
  
  # Add up all Wh cols
  #tempHhDT <- tempHhDT[, Sum := rowSums(.SD, na.rm = TRUE), .SDcols = grep("$", names(tempHhDT))] 
  
  hhStatTempDT <- tempHhDT[, .(nObs = .N),keyby = (date = as.Date(r_dateTime))] # can't do mean Wh as label varies
  hhStatTempDT <- hhStatTempDT[, hhID := hh]
  
  hhStatDT <- rbind(hhStatDT,hhStatTempDT) # add to the collector
  
  # > Save hh file ----
  ofile <- paste0(outPath, "1min/", hh,"_all_1min_data.csv")
  write_csv(tempHhDT, ofile)
  print(paste0("Saved ", ofile, ", gzipping..."))
  cmd <- paste0("gzip -f ", "'", path.expand(ofile), "'") # gzip it - use quotes in case of spaces in file name, expand path if needed
  try(system(cmd)) # in case it fails - if it does there will just be .csv files (not gzipped) - e.g. under windows
  print(paste0("Gzipped ", ofile))
}

#> Save observed data stats for all files loaded ----
ofile <- paste0(outPath, "hhDailyObservationsStats.csv")
print(paste0("Saving daily observations stats by hhid to ", ofile)) # write out version with file stats
write.csv(hhStatDT, ofile)
print("Done")
```

# Data quality analysis

Now produce some data quality plots & tables.

The following plots show the number of observations per day per household. In theory we should not see:

 * dates before 2014 or in to the future (they indicate data conversion errors)
 * more than 1440 observations per day (they indicate potentially duplicate data)

```{r loadedFilesObsPlots}
# tile plot ----
ggplot(hhStatDT, aes( x = date, y = hhID, fill = nObs)) +
  geom_tile() +
  scale_fill_gradient(low = "red", high = "green") +
  scale_x_date(date_labels = "%Y %b", date_breaks = "6 months") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + 
  labs(title = "N observations per household per day for all loaded grid spy data",
       caption = paste0(myCaption,
                        "\nOnly files of size > ", dataThreshold, " bytes loaded")
       
  )
ggsave(paste0(outPath, "gridSpyLoadedFileNobsTilePlot.png"))

# point plot ----
ggplot(hhStatDT, aes( x = date, y = nObs, colour = hhID)) +
  geom_point() +
  scale_x_date(date_labels = "%Y %b", date_breaks = "6 months") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + 
  labs(title = "N observations per household per day for all loaded grid spy data",
       caption = paste0(myCaption,
                        "\nOnly files of size > ", dataThreshold, " bytes loaded")
       
  )
ggsave(paste0(outPath, "gridSpyLoadedFileNobsPointPlot.png"))
```

The following table shows the min/max observations per day and min/max dates for each household. As above, we should not see:

 * dates before 2014 or in to the future (indicates date conversion errors)
 * more than 1440 observations per day (indicates potentially duplicate observations)
 
 We should also not see NA in any row (indicates date conversion errors). 
 
 If we do see any of these then we still have data cleaning work to do!

```{r summaryTable}
# Stats table (so we can pick out the dateTime errors)
t <- hhStatDT[, .(minObs = min(nObs),
             maxObs = max(nObs), # should not be more than 1440, if so suggests duplicates
             minDate = min(date),
             maxDate = max(date)),
         keyby = .(hhID)]

kable(caption = "Summary observation stats by hhID", t)
```

# Runtime


```{r check runtime}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r elapsed` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * ggplot2 - for slick graphics [@ggplot2]
 * dplyr - for select and contains [@dplyr]
 * lubridate - date manipulation [@lubridate]
 * knitr - to create this document [@knitr]
 * greenGridr - for local NZ GREEN Grid utilities
 
```{r sessionInfo}
sessionInfo()
```