---
title: "XXX"
author: "Ben Anderson (b.anderson@soton.ac.uk `@dataknut`)"
date: 'Last run at: `r Sys.time()`'
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    self_contained: no
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
bibliography: '`r paste0(findParentDirectory("GREENGrid"), "/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
options(knitr.table.format = 'markdown') # try to fix the tables issue (seems to be pushing html into latex)
```


```{r codeSetup, include=FALSE}
# Set start time ----
startTime <- proc.time()

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "dplyr", # data munching
             "ggplot2", # for fancy graphs
             "readr", # writing to files
             "skimr", # for skim
             "knitr" # for kable
)
# load them
loadLibraries(rmdLibs)

# run gg set up
nzGREENGrid::setup()

# Local paramaters
ggParams$dataLoc <- paste0(nzGREENGrid::findParentDirectory("nzGREENGrid"), "/data/")

ggParams$figCaption <- paste0("Source: June & December 1998 and June & December 2017 wholesale generation data",
                              "\nhttps://www.emi.ea.govt.nz/Wholesale/Datasets/Generation/Generation_MD/")

```

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r 1900 + as.POSIXlt(Sys.Date())$year`) _XXX_, `r ggParams$pubLoc`

This work is (c) `r as.POSIXlt(Sys.time())$year + 1900` the University of Southampton.

\newpage

# About

Report circulation:

 * Restricted to: [NZ GREEN Grid](https://www.otago.ac.nz/centre-sustainability/research/energy/otago050285.html) project partners and contractors.

## Purpose

This report is intended to: 

 * XYZ.

## Requirements:

 * ABC

## History

```{r child=ggParams$historyGenericRmd}
```
 
## Support

```{r child=ggParams$supportGenericRmd}
```
 

# Introduction

To ...

# Load data

# Analysis

# Discuss your results
here

# Conclusions
go here

# Runtime

```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * lubridate - date manipulation [@lubridate]
 * ggplot2 - for slick graphics [@ggplot2]
 * readr - for csv reading/writing [@readr]
 * dplyr - for select and contains [@dplyr]
 * progress - for progress bars [@progress]
 * knitr - to create this document & neat tables [@knitr]
 * nzGREENGrid - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
