---
title: 'FlipTheFleet Black Box Data Tests'
subtitle: 'Exploration of test data'
author: 'Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)'
date: 'Last run at: `r Sys.time()`'
output:
  html_document:
    code_folding: hide
    fig_caption: true
    keep_md: true
    number_sections: true
    self_contained: no
    toc: true
    toc_float: true
    toc_depth: 2
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
bibliography: '`r path.expand("~/bibliography.bib")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
options(knitr.table.format = 'markdown') # try to fix the tables issue (seems to be pushing html into latex)
```

```{r codeSetup, include=FALSE}
rm(list=ls(all=TRUE)) # remove all objects from workspace

# Set start time ----
startTime <- proc.time()

library(nzGREENGrid)

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "dplyr", # data munching
             "ggplot2", # for fancy graphs
             "readr", # writing to files
             "knitr" # for kable
)
# load them
nzGREENGrid::loadLibraries(rmdLibs)

# run gg set up
ggParams <- nzGREENGrid::setup()

# Local paramaters
ggParams$dataLoc <- paste0(nzGREENGrid::findParentDirectory("nzGREENGrid"), "/data/flipTheFleet/")
```

\newpage

# Citation

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r 1900 + as.POSIXlt(Sys.Date())$year`) FlipTheFleet Black Box Data Tests, `r pubLoc`.

This work is (c) `r as.POSIXlt(Sys.time())$year + 1900` the University of Southampton.

\newpage

# Introduction

Report circulation:

 * Restricted to: [NZ GREEN Grid](https://www.otago.ac.nz/centre-sustainability/research/energy/otago050285.html) project partners and contractors.

## Purpose

This report is intended to: 

 * load and test preliminary 'black box' EV monitoring data provided for assessment purposes by [FlipTheFleet](http://flipthefleet.org/).

## Requirements:

 * test dataset stored at XXX

## History

```{r include=ggParams$historyGenericRmd}
```
 
## Support

```{r include=ggParams$supportGenericRmd}
```
 

# Load data files

## EV test data

```{r set fileName}
ggParams$ftfFile <- paste0(ggParams$dataLoc, "EVBlackBox export 2018-06-10-233146.csv")
```

In this section we load and describe the  data files from `r ggParams$ftfFile`.

```{r load ftf data}
ftfDT <- data.table::as.data.table(readr::read_csv(ggParams$ftfFile))

Hmisc::describe(ftfDT)
```


 
# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * lubridate - date manipulation [@lubridate]
 * ggplot2 - for slick graphics [@ggplot2]
 * readr - for csv reading/writing [@readr]
 * dplyr - for select and contains [@dplyr]
 * progress - for progress bars [@progress]
 * knitr - to create this document & neat tables [@knitr]
 * nzGREENGrid - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
