---
params:
 title: "FlipTheFleet Test Black Box Data: Testing Time of Charging"
 subtitle: "Exploration of test data"
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: 'Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)'
date: 'Last run at: `r Sys.time()`'
# output:
#   html_document:
#     code_folding: hide
#     fig_caption: true
#     keep_md: true
#     number_sections: true
#     self_contained: no
#     toc: true
#     toc_float: true
#     toc_depth: 2
#   pdf_document:
#     fig_caption: yes
#     keep_tex: yes
#     number_sections: yes
#     toc: yes
#     toc_depth: 2
# bibliography: '`r path.expand("~/bibliography.bib")`'
output:
  bookdown::pdf_document2: default
  bookdown::html_document2:
    toc: true
    toc_float: true
bibliography: '`r paste0(nzGREENGrid::findParentDirectory("nzGREENGrid"), "/bibliography.bib")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
options(knitr.table.format = 'markdown') # try to fix the tables issue (seems to be pushing html into latex)
```

```{r codeSetup, include=FALSE}
rm(list=ls(all=TRUE)) # remove all objects from workspace

# Set start time ----
startTime <- proc.time()

library(nzGREENGrid)

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "dplyr", # data munching
             "ggplot2", # for fancy graphs
             "lubridate", # date & time processing
             "readr", # writing to files
             "knitr" # for kable
)
# load them
nzGREENGrid::loadLibraries(rmdLibs)

# run gg set up
nzGREENGrid::setup() # sets a load of parameters which are then accessible via ggParams$xxx

# Local paramaters
ggParams$dataLoc <- path.expand("/Volumes/hum-csafe/Research Projects/GREEN Grid/_RAW DATA/flipTheFleet/")
```

\newpage

# Citation

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r 1900 + as.POSIXlt(Sys.Date())$year`) `r params$title`: `r params$subtitle`, `r ggParams$pubLoc`.

This work is (c) `r as.POSIXlt(Sys.time())$year + 1900` the University of Southampton.

\newpage

# About

## Circulation

```{r generic circulation, child=ggParams$circulationGenericRmd}
```
 
## Purpose

This report is intended to: 

 * load and test preliminary 'black box' EV monitoring data provided for assessment purposes by [FlipTheFleet](http://flipthefleet.org/)
 * preliminary analysis of time of charging to understand impact on 'peak' electricity

## Requirements:

 * test dataset stored at `r ggParams$dataLoc`

## History

```{r child=ggParams$historyGenericRmd}
```
 
Specific history of this code:

 * https://git.soton.ac.uk/ba1e12/nzGREENGrid/tree/master/analysis/ev

## Support

```{r child=ggParams$supportGenericRmd}
```
 

# Load data files

## EV test data

```{r set fileName}
ggParams$ftfFile <- paste0(ggParams$dataLoc, "EVBlackBox export 2018-06-10-233146.csv")
```

In this section we load and describe the data from `r ggParams$ftfFile`. Note that we remove the following variables before we do so as they are potentially disclosive:

 * `Reg No`
 * `Latitude`
 * `Longitude` 
 * `Course (deg)` 

```{r load ftf data}
ftfDT <- data.table::as.data.table(readr::read_csv(ggParams$ftfFile))

# do analysis on safe version
ftfSafeDT <- createSafeFtF(ftfDT)
```

Create some useful derived variables.

```{r process ftf data, echo = TRUE}
# create deried 
ftfSafeDT <- createDerivedFtF(ftfSafeDT)
```

See ftFBlackBoxTestDataCodebook.Rmd for a full description of the data. 

Check charger related variables.

```{r checkCharging}
ggplot(ftfSafeDT, aes(x = `Charger (amps)`)) +
  geom_histogram()

ggplot(ftfSafeDT, aes(x = `Charger (V)`)) +
  geom_histogram()

ggplot(ftfSafeDT, aes(x = `Charger (amps)` , y = `Charger (V)` )) +
  geom_point()
```


# Timing of charging

We assume `Charger (amps)` is a good indicator of charging?

```{r plotChargingTime}
plotDT <- ftfSafeDT[, .(meanAmp = mean(`Charger (amps)`),
                        sumAmp = sum(`Charger (amps)`)), keyby = .(rTime)]

ggplot(plotDT, aes(x = rTime, y = meanAmp)) +
  geom_point()

ggplot(plotDT, aes(x = rTime, y = log(sumAmp))) +
  geom_point()
```

So this car appears to be on an overnight charging timer, although we do still see charging in the evening. We need a way to infer when the car is 'at home' without being disclosive. Perhaps we could use the modal overnight `Latitude` <-> `Longtidude` as 'home'?
# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * lubridate - date manipulation [@lubridate]
 * ggplot2 - for slick graphics [@ggplot2]
 * readr - for csv reading/writing [@readr]
 * dplyr - for select and contains [@dplyr]
 * progress - for progress bars [@progress]
 * knitr - to create this document & neat tables [@knitr]
 * nzGREENGrid - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
