---
title: "Misc lighting calculations"
author: "Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    toc: yes
    toc_depth: 2
    toc_float: TRUE
bibliography: '`r path.expand("~/bibliography.bib")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(GREENGridData)

library(data.table)
library(ggplot2)
library(hms)
library(kableExtra)
library(lubridate)
library(skimr)
```

# Get GREEN Grid Lighting Data

Get Carsten's re-weighted lighting profile data - derived from [@anderson_new_2018]

```{r getLight}
ggLightDT <- data.table::fread("~/Dropbox/Work/Otago_CfS_Ben/Students/CarstenDortans_MSc/Writing bursary/Lighting.csv")

skimr::skim(ggLightDT)

head(ggLightDT)
```

Check that the datetimes are NZ time otherwise we won't be comparing the same half-hours.

```{r checkTimes}
ggLightDT <- ggLightDT[, r_dateTime := lubridate::as_datetime(r_dateTime)]
ggLightDT$V1 <- NULL # how did that get there?
ggLightDT <- ggLightDT[, hms := hms::as.hms(r_dateTime)]
ggLightDT <- ggLightDT[, halfHour := hms::trunc_hms(hms, 60*30)]
                                              
ggLightDT <- GREENGridData::addNZSeason(ggLightDT)
plotDT <- ggLightDT[, .(meanGW = mean(powerW/10000000)), keyby = .(halfHour, season)]
p <- ggplot2::ggplot(plotDT, aes(x = halfHour, y = meanGW, colour = season)) +
  geom_line() +
  labs(x = "Time of Day",
       y = "Mean power demand (Lighting, GW)",
       caption = "GREEN Grid lighting data re-weighted to NZ population")
p
```

Looks OK. This is because `hms::as.hms()` cleverly converts `r_dateTime` which it assumes to be UTC (it is) into local time. Without telling us...

Now convert that to half-hourly GWh for comparison with EA data.

```{r convertToHHGwh}
# do this before aggregating
ggLightDT <- ggLightDT[, powerWh := powerW/60] # cos it's per minute
ggLightDT <- ggLightDT[, consumptionGWh := powerWh/1000000000]

# need to aggregate from 1 min to (all) half hours
ggLightDT <- ggLightDT[, halfHourDate := lubridate::floor_date(r_dateTime, "30 mins")]
dt <- ggLightDT[, .(consumptionGWh = sum(consumptionGWh)), keyby = .(halfHourDate, halfHour, season)]

# now get the mean
plotGG_DT <- dt[, .(MeanGWh = mean(consumptionGWh)), keyby = .(halfHour, season)]

# check
message("Sum of GWh before agg = ", sum(ggLightDT$consumptionGWh))
message("Sum of GWh after agg = ", sum(dt$consumptionGWh))

p <- ggplot2::ggplot(plotGG_DT, aes(x = halfHour, y = MeanGWh, colour = season)) +
  geom_line() +
  labs(x = "Time of Day",
       y = "Mean energy consumption (Lighting, GWh)",
       caption = "GREEN Grid lighting data re-weighted to NZ population")
p

```

# Get EA Generation data for 2015

Get the 2015 Generation data from the EA (igores distributed gen - solar as not really relevant to evening peaks). Data pre-downloaded from https://www.emi.ea.govt.nz/Wholesale/Datasets/Generation/Generation_MD & cleaned.

```{r getGen}
eaPath <- "~/Data/NZ_GREENGrid/ea/"

fList <- data.table::as.data.table(list.files(eaPath))

f2015 <- fList[V1 %like% "2015" & V1 %like% "long"]
f2015 <- f2015[, fullName := paste0(eaPath, V1)]
ea2015GenDT <- data.table::data.table() # data bucket

# load them all in one go very fast
ea2015GenDT = rbindlist(lapply(f2015$fullName, fread))

skimr::skim(ea2015GenDT)

head(ea2015GenDT)
```

Now do the same process of checking and aggregating.

```{r checkGen}
ea2015GenDT <- ea2015GenDT[, r_dateTime := lubridate::as_datetime(rDateTime)]
ea2015GenDT <- ea2015GenDT[, r_dateTime := lubridate::force_tz(r_dateTime, tzone = "Pacific/Auckland")] # got to force it

ea2015GenDT <- ea2015GenDT[, hms := hms::as.hms(r_dateTime)]
ea2015GenDT <- ea2015GenDT[, halfHour := hms::trunc_hms(hms, 60*30)]

ea2015GenDT <- GREENGridData::addNZSeason(ea2015GenDT)

# make GWh
ea2015GenDT <- ea2015GenDT[, consumptionGWh := kWh/1000000]

# add up to total GWh for eachhalf hour (currently it's per gen site/fuel)
aggea2015GenDT <- ea2015GenDT[, .(consumptionGWh = sum(consumptionGWh)), keyby = .(season, r_dateTime, halfHour)]

# check
message("Sum of GWh = ", sum(aggea2015GenDT$consumptionGWh))

# now re-create plotDT for the mean
plotEA_DT <- aggea2015GenDT[, .(MeanGWh = mean(consumptionGWh)), keyby = .(halfHour, season)]
p <- ggplot2::ggplot(plotEA_DT, aes(x = halfHour, y = MeanGWh, colour = season)) +
  geom_line() +
  labs(x = "Time of Day",
       y = "Mean energy consumption per half-hour (All, GWh)",
       caption = "EA Wholesale Generation data (excl. distributed solar)")
p

```

# Lighting as a % of total consumption

So now we need to plot the contribution of lighting to this.

```{r combineData}
setkey(plotEA_DT, season, halfHour)
setkey(plotGG_DT, season, halfHour)

plotEA_DT <- plotEA_DT[, eaMeanGWh := MeanGWh]
plotGG_DT <- plotGG_DT[, lightingMeanGWh := MeanGWh]

plotDT <- plotEA_DT[plotGG_DT]
plotDT <- plotDT[, pc_lighting := 100*(lightingMeanGWh/eaMeanGWh)]

p <- ggplot2::ggplot(plotDT, aes(x = halfHour, y = pc_lighting, colour = season)) +
  geom_line() +
  labs(x = "Time of Day",
       y = "Lighting as a % of total generation (GWh)",
       caption = "EA Wholesale generation data & GREEN Grid population weighted lighting data")
  
p

t <- summary(plotDT[, .(eaMeanGWh,lightingMeanGWh,pc_lighting )])

kableExtra::kable(t, caption = "Summary of data results") %>%
  kable_styling()
```

```{r theEnd}

```

# R packages used

 * GREENGridData - [@GREENGridData]
 * data.table - [@data.table]
 * ggplot2 - [@ggplot2]
 * hms - [@hms]
 * kableExtra - [@kableExtra]
 * lubridate - [@lubridate]
 * skimr - [@skimr]

# References
