---
params:
 title: "Technical Potential of Demand Response"
 subtitle: "Heat Pump Analysis"
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: 'Carsten Dortans (xxx@otago.ac.nz)'
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    toc: true
    toc_float: TRUE
    toc_depth: 2
    keep_md: TRUE
    self_contained: no
  bookdown::pdf_document2:
    toc: true
    toc_depth: 2
bibliography: '`r paste0(nzGREENGrid::findParentDirectory("nzGREENGrid"), "/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # by default turn off code echo
#options(knitr.table.format = 'markdown') # try to fix the tables issue (seems to be pushing html into latex)
```

```{r codeSetup, include=FALSE}
#rm(list=ls(all=TRUE)) # remove all objects from workspace

# Set start time ----
startTime <- proc.time()

library(nzGREENGrid)

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "dplyr", # data munching
             "ggplot2", # for fancy graphs
             "lubridate", # date & time processing
             "readr", # writing to files
             "skimr", # for skim
             "knitr" # for kable
)
# load them
nzGREENGrid::loadLibraries(rmdLibs)

# run gg set up
nzGREENGrid::setup() # sets a load of parameters which are then accessible via ggParams$xxx

# Local paramaters

sysInfo <- Sys.info()

userName <- sysInfo[[7]]
if(userName == "ben"){
  ggParams$dataLoc <- path.expand("~/Dropbox/Work/Carsten_MSc/ggData/profiles/")
} 
if(userName == "carsten.dortans"){
  ggParams$dataLoc <- path.expand("~/Dropbox/Carsten_MA/ggData/profiles/")
}
```

\newpage

# Citation

If you wish to use any of the material from this report please cite as:

 * Dortans, C. (`r 1900 + as.POSIXlt(Sys.Date())$year`) `r params$title`: `r params$subtitle`, `r ggParams$pubLoc`.

This work is (c) `r as.POSIXlt(Sys.time())$year + 1900` the University of Southampton.

\newpage

# About

## Circulation

```{r generic circulation, child=ggParams$circulationGenericRmd}
```
 
## Purpose

This report is intended to: 

 * load and test GREEN Grid heat pump and hot water profiles.

## Requirements:

 * test dataset stored at `r ggParams$dataLoc`

## History

```{r child=ggParams$historyGenericRmd}
```
 
Specific history of this code:

 * https://git.soton.ac.uk/ba1e12/nzGREENGrid/tree/master/analysis/demandResponse

## Support

```{r child=ggParams$supportGenericRmd}
```
 

# Load data files

## Heat pump profiles

This file is the pre-aggregated data for all heat pump circuits in the GREEN Grid data for April 2015 - March 2016 (check!)

```{r set fileName}
ggParams$profilesFile <- paste0(ggParams$dataLoc, "Heat Pump_2015-04-01_2016-03-31_overallSeasonalProfiles.csv.gz")
```

In this section we load and describe the  data files from `r ggParams$profilesFile`.

```{r load data}
print(paste0("Trying to load: ", ggParams$profilesFile))
heatPumpProfileDT <- data.table::as.data.table(readr::read_csv(ggParams$profilesFile))
```


Describe using skim:

```{r skimRable}
skimr::skim(heatPumpProfileDT)
```

Draw a plot of GreenGrid heat pump profiles.

```{r profilePlot, fig.cap="Heat pump profiles"}
myPlot <- ggplot2::ggplot(heatPumpProfileDT, aes(x = obsHourMin, colour = season)) +
  geom_point(aes(y = meanW)) +
  facet_grid(season ~ .)

myPlot
```
#Scaling method 1

Now draw a plot of what woud happen if we scaled this up to all NZ households?

Figure \@ref(fig:scaledUpPlots)

```{r scaledUpPlots,fig.cap='Mean Load Heat Pumps by Season'}
nzHH <- 1549890

heatPumpProfileDT <- heatPumpProfileDT[, scaledMW := (meanW * nzHH)/10^6]

myPlot <- ggplot2::ggplot(heatPumpProfileDT, aes(x = obsHourMin, colour = season)) +
  geom_point(aes(y = scaledMW)) +
  facet_grid(season ~ .)

myPlot
```
#Scaling method 2
Alternative calculation method: Assuming EECA data is correct for heat pump value, 1) generating the percentage of total load (peroftotal) while telling data.table to create a new column with the calculation of the percentage. We then multiplied EECA's total GWh with the percentage
```{r new calc}
totalGWH<-708

summeanW<-heatPumpProfileDT[,sum(meanW)]



heatPumpProfileDT <- heatPumpProfileDT[, EECApm := (meanW / summeanW) * totalGWH] 

myPlot <- ggplot2::ggplot(heatPumpProfileDT, aes(x = obsHourMin, colour = season)) +
  geom_point(aes(y = EECApm)) +
  facet_grid(season ~ .) +
  labs(x='Time of Day', y='GWh')

myPlot
```



# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * lubridate - date manipulation [@lubridate]
 * ggplot2 - for slick graphics [@ggplot2]
 * readr - for csv reading/writing [@readr]
 * skimr - for skim [@skimr]
 * knitr - to create this document & neat tables [@knitr]
 * nzGREENGrid - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
