---
params:
  hhID: "" 
  title: 'NZ GREEN Grid project example:'
  subtitle: 'Testing power demand: '
title: '`r params$title`'
subtitle: '`r paste0(params$subtitle, params$hhID)`'
author: "Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    code_folding: hide
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    self_contained: no
    toc: yes
    toc_depth: 2
    toc_float: TRUE
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
bibliography: '`r paste0(findParentDirectory("GREENGrid"), "/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
options(knitr.table.format = 'markdown') # try to fix the tables issue (seems to be pushing html into latex)
```


```{r codeSetup, include=FALSE}

# Set start time ----
startTime <- proc.time()

# Load nzGREENGrid package ----
library(GREENGrid) # local utilities

GREENGrid::setup()

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "lubridate", # date & time stuff
             "ggplot2", # for fancy graphs
             "plotly", # interactive charts - good for data checking
             "readr", # for reading & parsing .csv files
             "hms", # for hms coding
             "kableExtra", # for fancy tables
             "knitr" # for kable
)
# load them
loadLibraries(rmdLibs)

# Local parameters ----
# change this to suit your data location & to use a different household
gsFile <- paste0(ggParams$dataLoc, "Clean_data/safe/gridSpy/1min/data/", params$hhID,"_all_1min_data.csv.gz")
hhFile <- paste0(ggParams$dataLoc, "Clean_data/safe/survey/ggHouseholdAttributes.csv")

# Local functions ----


```

\newpage

# About

## Report circulation:

 * Public - this report is intended to accompany the data release.
 
## License

```{r ccby license, child=ggParams$licenseCCBY}
```
 
## Citation

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r lubridate::year(today())`) `r params$title` `r paste0(params$subtitle, params$hhID)` `r ggParams$pubLoc`.

This work is (c) `r lubridate::year(today())` the University of Southampton.

## History

 * [Report history](hhttps://github.com/CfSOtago/GREENGrid/tree/master/dataProcessing/gridSpy)
 
## Requirements:

This report uses the safe version of the grid spy 1 minute data which has been processed using https://github.com/CfSOtago/GREENGridData/tree/master/dataProcessing/gridSpy.

## Support

```{r generic support, child=ggParams$supportGeneric}
```
 
\newpage

# Introduction

```{r generic sample, child=ggParams$sampleGeneric}
```
 
This report provides summary analysis of one household as an example.

# Load `r params$hhID` data

The data used to generate this report is:

 * `r gsFile`
 * `r hhFile`

```{r hhData}
hhDT <- data.table::as.data.table(readr::read_csv(hhFile)) # load hh data

t <- hhDT[hhID == params$hhID]

kableExtra::kable(t, 
             booktabs = TRUE,
             caption = paste0("Summary of household data for", params$hhID)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), font_size = 9)

# Guide to markdown cross-refs: https://bookdown.org/yihui/rmarkdown/bookdown-markdown.html#bookdown-markdown
```

Table \@ref(tab:hhData) shows household attributes such as how many people live in this household.

Table \@ref(tab:gsData) shows the initial grid spy 1 minute power data. 

```{r gsData, fig.cap="Example of grid spy data loaded"}
gsDT <- data.table::as.data.table(readr::read_csv(gsFile, 
                                                  col_types = cols(hhID = col_character(),
                                                                   dateTime_orig = col_character(), # <- this is crucial otherwise readr attempts to parse this as a dateTime and FAILS (see https://github.com/CfSOtago/GREENGridData/issues/22)
                                                                   TZ_orig = col_character(),
                                                                   r_dateTime = col_datetime(format = ""),
                                                                   circuit = col_character(),
                                                                   powerW = col_double() # <- also crucial otherwise readr seems to assume an integer
                                                                   )
                                                  )
                                  ) # load gs data
t <- head(gsDT)
kableExtra::kable(t, caption = paste0("Summary of grid spy data for", params$hhID)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), font_size = 9)
```

Note that:

 * the original dataTime (`dateTime_orig`) and TZ (`TZ_orig`) have been retained so that the user can check for parsing errors (see https://github.com/CfSOtago/GREENGridData/issues/2) if required;
 * r_dateTime loads as UTC by default so we need to re-timezone it to NZ time before deriving any other times and dates.

```{r gsDataFix, fig.cap="Summary of grid spy data loaded", echo=TRUE}
# r_dateTime will have loaded as UTC (if you don't believe me, check :-)
gsDT <- gsDT[, r_dateTime := with_tz(r_dateTime, tzone = "Pacific/Auckland")]
t <- head(gsDT)
gsDT <- gsDT[, obsDate := lubridate::date(r_dateTime)]
gsDT <- gsDT[, obsTime := hms::as.hms(r_dateTime)]
kableExtra::kable(t, caption = paste0("Summary of grid spy data for", params$hhID, "tz corrected")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), font_size = 9)

gsDT <- gsDT[, month := lubridate::month(r_dateTime, label = TRUE)]
gsDT <- gsDT[, year := lubridate::year(r_dateTime)]


```


# Plot monthly power profiles

Figure \@ref(fig:plotProfiles) shows plots overall mean power (W) per minute per month for each circuit to show:

 * patterns of missing data (no lines)
 * patterns of consumption
 * possible dateTime issues (where consumption patterns seem to be stangely shifted in time)
 * possible PV installation

```{r plotProfiles, fig.height=8}
plotDT <- gsDT[, .(meanW = mean(powerW)), keyby = .(circuit, month, year, obsTime)
             ] # aggregate by circuit to preserve unique circuit labels in households 
# (e.g. rf_46) where names are re-used but with different ids. see ?fixCircuitLabels_rf_46
vLineAlpha <- 0.4
vLineCol <- "#0072B2" # http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette
timeBreaks <- c(hms::as.hms("04:00:00"), 
                hms::as.hms("08:00:00"),
                hms::as.hms("12:00:00"),
                hms::as.hms("16:00:00"),
                hms::as.hms("20:00:00"),
                hms::as.hms("24:00:00")
)

myPlot <- ggplot2::ggplot(plotDT, aes(x = obsTime, y = meanW/1000, colour = circuit)) +
  geom_line() + 
  facet_grid(month  ~ year) + 
  theme(strip.text.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + 
  theme(legend.position = "bottom") + 
  labs(title = paste0("Montly mean power profiles by circuit plot: ", params$hhID),
       y = "Mean kW", 
       caption = paste0("gridSpy data from ", min(gsDT$r_dateTime), 
                        " to ", max(gsDT$r_dateTime),
                        "\nobsTime = Pacific/Auckland"))

myPlot + 
  scale_x_time(breaks = timeBreaks) +
  geom_vline(xintercept = timeBreaks, alpha = vLineAlpha, colour = vLineCol)
```

Figure \@ref(fig:plotProfiles) in rf_38 seems to indicate some potential time of day mis-alignments in late 2016. We suspect these may be due to incorrect date formats.

Figure \@ref(fig:dateCheck) checks the original date format by month and year.

```{r dateCheck}
plotDT <- gsDT[, .(nObs = .N), keyby = .(TZ_orig, month, year, obsTime)]
ggplot(plotDT, aes(x = obsTime, colour = TZ_orig, y = nObs)) + 
  geom_line() + 
  facet_grid(month ~ as.factor(year)) +
  theme(strip.text.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))
```

Figure \@ref(fig:plotlyProfiles) repeats the original power demand analysis but just for the few 'doubtful' months so that we can check and also compare with another household where we do not see this (e.g. rf_06). It uses plotly to generate an interactive plot (html only).

```{r selectMonth}
testDT <- gsDT[month == "Oct" & year == 2016 | month == "Nov" & year == 2016 |
                 month == "Dec" & year == 2016 | 
                 month == "Jan" & year == 2017]

t <- head(testDT[obsTime > hms::as.hms("02:00:00")], 10)

kableExtra::kable(caption = "First 10 lines of selected data", t) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), font_size = 9)
```

```{r plotlyProfiles, fig.height=8}
plotDT <- testDT[, .(meanW = mean(powerW)), keyby = .(circuit, month, year, obsTime)
             ] # aggregate by circuit to preserve unique circuit labels in households 
# (e.g. rf_46) where names are re-used but with different ids. see ?fixCircuitLabels_rf_46

myPlot <- ggplot2::ggplot(plotDT, aes(x = obsTime, y = meanW/1000, colour = circuit)) +
  geom_line() + 
  facet_grid(month  ~ year) + 
  theme(strip.text.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + 
  theme(legend.position = "bottom") + 
  labs(title = paste0("Montly mean power profiles by circuit plot: ", params$hhID),
       y = "Mean kW", 
       caption = paste0("gridSpy data from ", min(gsDT$r_dateTime), 
                        " to ", max(gsDT$r_dateTime),
                        "\nobsTime = Pacific/Auckland"))

myPLot <- myPlot + 
  scale_x_time(breaks = timeBreaks) +
  geom_vline(xintercept = timeBreaks, alpha = vLineAlpha, colour = vLineCol)

plotly::ggplotly(myPlot)
```


Figure \@ref(fig:checkDoubtfulDates) repeats the earlier analysis but shows just the 'doubtful' months. If the plots do not show multiple date formats then it seems unlikely there is an issue.


```{r checkDoubtfulDates}
plotDT <- testDT[, .(nObs = .N), keyby = .(TZ_orig, month, year, obsTime)]
ggplot(plotDT, aes(x = obsTime, colour = TZ_orig, y = nObs)) + 
  geom_line() + 
  facet_grid(month ~ as.factor(year)) +
  theme(strip.text.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))
```

# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * lubridate - date manipulation [@lubridate]
 * ggplot2 - for slick graphics [@ggplot2]
 * readr - for csv reading/writing [@readr]
 * knitr - to create this document & neat tables [@knitr]
 * GREENGrid - for local NZ GREEN Grid project utilities
 * GREENGridData - for local NZ GREEN Grid project data utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
